---
title: "DS311 - R Lab Assignment"
author: "Jonathan Terada"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
ncol(mtcars)
nrow(mtcars)
# Answer:
print(paste("There are total of", ncol(mtcars), "variables and", nrow(mtcars), "observations in this data set."))

```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)

# Answer:
print("There are 7 discrete variables and 4 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m <- mean(mtcars$mpg)
v <- var(mtcars$mpg)
s <- sd(mtcars$mpg)

print(paste("The average of Mile Per Gallon from this data set is ", m, " with variance ",  v, " and standard deviation",  s, "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
avg_mpg_cyl <- aggregate(mpg ~ cyl, data = mtcars, mean)
avg_mpg_cyl

sd_mpg_gear <- aggregate(mpg ~ gear, data = mtcars, sd)
sd_mpg_gear

 ```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
cylinder_obs <- table(mtcars$cyl, mtcars$gear)
cylinder_obs

print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
plot(weight ~ group, data = PlantGrowth,
     main = "Plant Growth by Group",
     xlab = "Group",
     ylab = "Weight")
summary(PlantGrowth)

```

Result:

=> Report a paragraph to summarize your findings from the plot!

I found that the control group (ctrl) had a median plant growth of 5.1, with a range of values from a minimum of 3.5 to a maximum of 6.3. In comparison, the first treatment group (trt1) had a median plant growth of 4.5, with a range of values from 3.6 to 5.8. The second treatment group (trt2) had a higher mean plant growth of 5.5, with a range of values from 5.0 to 6.4. Overall, based on these statistics, it appears that the control group had a better median, minimum, and maximum plant growth compared to the other two treatment groups.


b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mtcars$mpg, breaks = 10,
     main = "Histogram of MPG",
     xlab = "Miles per gallon",
     ylab = "Num of Cars")


print("Most of the cars in this data set are in the class of 15 mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
pairs(USArrests, main = "Pairs Plot of USArrests Data")

plot(USArrests$Murder, USArrests$Assault, 
     main = "Scatter Plot of Murder and Assault", 
     xlab = "Murder", 
     ylab = "Assault")

```

Result:

=> Report a paragraph to summarize your findings from the plot!

I observed a wide range of patterns in the data related to arrests. It was fascinating to see how each plot displayed a unique relationship between the variables being examined.

In particular, when I analyzed the plot of arrests against murders, I observed that the data points were scattered across the plot, with no clear pattern emerging. However, I did note a strong positive correlation between the two variables, with both increasing together in most cases.

***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
summary(housingData)
summary(housingData$Year.Built)
summary(housingData$Market.Value.per.SqFt)

m <- mean(housingData$Market.Value.per.SqFt)
v <- var(housingData$Market.Value.per.SqFt)
s <- sd(housingData$Market.Value.per.SqFt)

max(housingData$Market.Value.per.SqFt)
min(housingData$Market.Value.per.SqFt)

avg_MVPSqFt_Year <- aggregate(Market.Value.per.SqFt ~ Year.Built, data = housingData, mean)
avg_MVPSqFt_Year

avg_MVPSqFt_Boro <- aggregate(Market.Value.per.SqFt ~ Boro, data = housingData, mean)
avg_MVPSqFt_Boro

avg_MVPSqFt_Fin <- aggregate(Market.Value.per.SqFt ~ Neighborhood, data = housingData, mean)
avg_MVPSqFt_Fin

sd__MVPSqFt_Year <- aggregate(Market.Value.per.SqFt ~ Year.Built, data = housingData, sd)
sd__MVPSqFt_Year

var__MVPSqFt_Year <- aggregate(Market.Value.per.SqFt ~ Year.Built, data = housingData, var)
var__MVPSqFt_Year
```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
hist(housingData$Market.Value.per.SqFt, breaks = 10,
     main = "Histogram of Market Value per SqFt", 
     xlab = "Market Value per SqFt", 
     ylab = "Frequency")
plot(avg_MVPSqFt_Year$Year.Built, avg_MVPSqFt_Year$Market.Value.per.SqFt,
     main = "Scatter Plot of Market Value per SqFt and the years", 
     xlab = "Year", 
     ylab = "Market Value per SqFt")

```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!

In this exercise, I analyzed housing data and discovered interesting insights. The market value per SqFt ranged from 10.66 to 399.38, with an average of 133.17 and a standard deviation of 68.88. The scatter plot of market value per SqFt and year of construction showed no clear correlation, but data points clustered around 100-200 market value per SqFt between 1900 and 2000. Overall, this exercise was a great learning experience.
